---
title: "Introduction to Shiny"
author: "Joris Meys"
date: "05/09/2019"
runtime: shiny
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Controlling reactivity

## Exercise

Reactivity will be demonstrated by an exercise.

During this class, we'll develop an app that :

 - lets the user upload a file & shows the data
 - lets the user select a X and Y variable
 - shows the distribution of both X and Y
 - fits a linear model Y ~ X and shows the results
 - lets the user download the predictions for Y
 
## Working with files

```{r, echo = TRUE}
fileInput("afile","Select a file",
          accept = "text/csv")
```

```{r, echo = TRUE}
renderPrint({
  str(input$afile)
})
```

## Uploading files

Use `filInput()`.

```{r, eval = FALSE, echo = TRUE}
fileInput("afile","Select a file",
          accept = "text/csv")
```

This returns a data frame with following info:

 - **name**: The original name of the file
 - **size**: The size of the file
 - **type**: The [MIME type](https://developer.mozilla.org/en-US/docs/Web/HTTP/Basics_of_HTTP/MIME_types) of the file **if** the browser could figure it out.
 - **datapath**: the actual location of the temporary file on the server. 
 

## Uploading files

To use the file, you need to use the `datapath` value!

```{r, echo = TRUE, eval=FALSE}
output$tbl <- renderDataTable({
  read.csv(input$afile$datapath)
})
```

**Sidenote**

The function `renderDataTable` from `shiny` can be extended using the `DT` package. The function `renderDT()` allows for
editable tables where observations can be selected.




## Exercise

Start by:

 - create a 2-file shiny app
 - setting up a page with a sidebar
 - add an upload widget for a file (make the data a reactive value)
 - add a tab with the data table of the uploaded data
 
## Controlling reactivity : only run with input

Problem: When there's no data file uploaded, the data table can't be displayed. 

Solution: `req()` (short for require)

 - tells the observer/reactive expression to only run for `truthy` values.
 - prevents error messages due to missing input
 
A `truthy` value is :

 - a loose, intuitive concept of `TRUE`
 - anything that's not `NULL`, `FALSE`, `""`, `NA`, ...
 - an inclicked `actionButton()`
 
** The value `0` is truthy, even though `isTRUE(0)` is `FALSE`.

## Create a dynamic UI

```{r, echo = TRUE}
sliderInput("max","Select maximum", min = 5, max=10,
            value = 5)
uiOutput("ui")
```

```{r, echo = TRUE}
output$ui <- renderUI({
  tagList(
    sliderInput("val","Select value",min = 1, 
                max = input$max, value = 1 )
  )
})
```

## Create a dynamic UI : Basics

The function `renderUI`:

 - should return a `tagList` with UI elements (input or output)
 - Is rendered on the UI side using `uiOutput()`.
 - Argument `outputArgs` allows to pass extra arguments to 
 `uiOutput()` if necessary.
 
If output of `renderUI` is `NULL`, `uiOutput` doesn't show anything.

## Exercise

Add the following to the app:

 - A UI with 2 `selectInput`s that let you select a X and a Y variable
 - Add two tabs with a histogram and summary values for
 the variables X and Y
 
## Checking the input with validate

Problem: If you select a factor variable, the histogram won't calculate and an error is shown

Solution: only run the code if the input is numeric with `validate()` and `need()`

 - `validate` takes one or more conditions to test.
 - conditions are evaluated with `need`
 - `need` allows for specific error message
 
```{r, echo = TRUE, eval = FALSE}
renderTable({
  validate(need(is.numeric(iris[[input$var1]]),
                "Please select a numeric variable."),
           need(is.factor(iris[[input$var2]]),
                "Var 2 needs to be a factor."))
})
```
 
## Controlling reactivity : action buttons

Problem : the model runs each time one of the variables is
changed. This is a waste of resources.

Solution: Control actions using an `actionButton()`.

```{r, echo = TRUE}
actionButton("Go","Click me!")
```

gives:

```{r, echo = TRUE}
renderText({paste("Button is clicked",input$Go,"times")})
```

## Controlling reactivity : Isolation

You want the expressive reaction to only run when the button is clicked. This means removing all other dependencies.

 - using `isolate()`: removes dependency from an input

```{r, echo = TRUE, eval = FALSE}
renderData({
  input$Go  # The button, but output is not used
  sel <- isolate(input$var) # The actual input, dependency removed
  data <- iris[[sel]]
})
```

 - using `eventReactive`: removes all dependencies except the reactive event
 
```{r, echo = TRUE, eval = FALSE}
thedata <- eventReactive(input$Go,{
  data <- iris[[sel]]
})
renderDataTable({thedata()})
``` 

## Exercise

 - Add the necessary controls so the histograms and the model are only built when both X and Y are numeric.
 - An extra tab `Model` that shows a graph of the model and 
 the model output (use either `verbatimTextOutput` or try the
 `broom` package for tidy output.)
 - Add an `actionButton` to the app in such a way that the model is only calculated when the action button is clicked.


## Downloading files

This can be done with the `downloadButton()` and `downloadHandler()` functions:

**UI**
```{r, echo = TRUE, eval = FALSE}
downloadButton("GetData","Download")
```

**server**
```{r, echo = TRUE, eval = FALSE}
output$GetData <- downloadHandler(
  filename = "example.csv",
  content = function(file){
    write.csv(iris, file, row.names = FALSE)
  }
)
```

## Exercise

Add a download button that returns a CSV file with :

 - the X variable with its original name
 - the Y variable with its original name
 - a column called `predict` with predictions.
 
Add a download button that lets the user download the plot of the model.

## Extra sources

 - Reactivity in full : http://shiny.rstudio.com/articles/reactivity-overview.html
 - use of isolate : http://shiny.rstudio.com/articles/isolation.html
 - scheduling of execution : http://shiny.rstudio.com/articles/execution-scheduling.html
 - Dynamic UI : https://shiny.rstudio.com/articles/dynamic-ui.html
 - Action buttons : http://shiny.rstudio.com/articles/action-buttons.html
 
 - Upload data: https://shiny.rstudio.com/articles/upload.html
 - Download data: https://shiny.rstudio.com/articles/download.html
 
 - DT package : https://rstudio.github.io/DT/shiny.html
 
