---
title: "Introduction to Shiny"
author: "Joris Meys"
date: "06/09/2021"
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval=FALSE)
```

# Layout styles

## Layout principles of apps

Let's recap:

 - **Page**: the main container of an app, contains panels.
     * `fluidPage`
     * `shinydashboard`
 - **Panel** : a container containing other panels or widgets.
     * `sidebarPanel`
     * `mainPanel`
 - **Widget**: a functional unit (input, output)
     * `sliderInput`
     * `plotOutput`

This general principle goes for all different layout designs and possibilities.

## Fluid layout using `fluidPage()`

Fluid design is the standard in shiny. This means :

 - row based: a row is created using `fluidRow()`
     * They don't have a fixed height!
     * Elements are placed on same line **if possible** (eg columns)!
     * Elements like input/output widgets are often forced to stack vertically.
 - Rows can be horizontally divided in columns using `column()`
     * there's 12 column units in a row
     * width of column in column units (1 - 12)

These are **panels** : can be nested in eachother

## Add HTML tags for formatting etc

`Shiny` contains a set of HTML functions :

 - `p()`: a paragraph with text
 - `br()`: a break
 - `a()`: a hyperlink
 - `h1()` to `h6()`: headers at different levels.
 
110 tag functions are contained in `tags` (See `ExampleLayout2`)
 
You can add attributes, eg:

```{r, echo=TRUE}
a("HTML tag reference!", href = "https://www.w3schools.com/TAGs/")
```

## Exercises

 - Replicate the design in the image for the app LinearRegression
 - Replicate the design in the image for the app BoxPlots
 

## Add navigation to the app

You can separate widgets in tabs using `tabPanel()`

Tabs have to be contained in :

 - `tabSetPanel()` : gives clickable tabs
 - `navlistPanel()` : gives navigation menu on the side
 - `navbarPage()` : a **page** with navigation bar for tabs
 
Keep in mind: `tabPanel()` is a container like `mainPanel()` etc.

## Control widget size :

Width and height can be set in CSS units for every widget.

 - is a character value with a number plus suffix.
 - suffices : `%`, `px`, `cm`, `mm`, `pt`, `in`, ...
 - relative width works relative to the column (eg `100%` = full column width)
 - relative height does NOT work! (height is determined by highest element.)
 
To use relative heights : 
 - use a `fillPage`
 - use a `fillRow` or `fillCol` with fixed height
 
**Widgets are contained in a specific class that forces them to stack vertically by default.**
**To place them next to eachother, either use columns or `splitLayout()`.**

## Exercises

Revisit the covid app and add:

 - A heatmap, controlled by the same inputs (region, gender and date range), next to the plot that's there. 
 - An extra tab with a summary table containing the average infections over the selected period for all regions. It's only controlled by gender and date range.
 
## Extra sources

 - Tab sets and fluid system : http://shiny.rstudio.com/articles/layout-guide.html
 - shiny HTML tags : http://shiny.rstudio.com/articles/html-tags.html
 - shiny HTML glossary : http://shiny.rstudio.com/articles/tag-glossary.html
 - Tutorial video : https://vimeo.com/rstudioinc/review/131218530/212d8a5a7a/#t=1h52m52s
 

# Running apps

## App structures

We've seen different kinds of apps :

 - **embedded in RMarkdown slides**
    * Only widgets needed
    * frontend/backend done for you
 - **single file apps in a file `app.R`**
    * Both UI and server in a single file
    * must return app object using `shinyApp()`
 - **two file apps: `ui.R` and `server.R`**
    * Must be in its own folder
    * `ui.R` returns a UI page object
    * `server.R` returns a function 

## Running an app.

Simplest way: clicking on a button in RStudio

More complex :

 - create `UI` and `server` objects
 - combine in a app object with `shinyApp()`
 - use `runApp()` to use the app.
 
```{r, eval = FALSE, echo = TRUE}
UIobj <- fluidPage( ... )
serverfun <- function(input, output){ ... }

theApp <- shinyApp(UIobj, serverfun)
runApp(theApp)
```

## Running an app.

Other ways to use `runApp()` :

 - using a list with a UI and server object:
```{r, eval = FALSE, echo = TRUE}
runApp(list(UI,serverfun))
```
 - Passing the directory where the files are located:
```{r, eval = FALSE, echo = TRUE}
runApp("path/to/appDir")
```

Allows us to create a function to call the app!

*See example `HistWithDens.R`*

## Exercise

Split the covid app in a ui.R and server.R, and try to run it from the command line.