---
title: "Introduction to Shiny"
author: "Joris Meys"
date: "05/09/2019"
runtime: shiny
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Layout styles

## Recap on app structure

We've seen different kinds of apps :

 - **embedded in RMarkdown slides**
    * Only widgets needed
    * frontend/backend done for you
 - **single file apps in a file `app.R`**
    * Both UI and server in a single file
    * must return app object using `shinyApp()`
 - **two file apps: `ui.R` and `server.R`**
    * Must be in its own folder
    * `ui.R` returns a UI page object
    * `server.R` returns a function 

## Running an app.

Simplest way: clicking on a button in RStudio

More complex :

 - create `UI` and `server` objects
 - combine in a app object with `shinyApp()`
 - use `runApp()` to use the app.
 
```{r, eval = FALSE, echo = TRUE}
UIobj <- fluidPage( ... )
serverfun <- function(input, output){ ... }

theApp <- shinyApp(UIobj, serverfun)
runApp(theApp)
```

## Running an app.

Other ways to use `runApp()` :

 - using a list with a UI and server object:
```{r, eval = FALSE, echo = TRUE}
runApp(list(UI,serverfun))
```
 - Passing the directory where the files are located:
```{r, eval = FALSE, echo = TRUE}
runApp("path/to/appDir")
```

Allows us to create a function to call the app!

*See example `HistWithDens.R`*