---
title: "Introduction to Shiny"
author: "Joris Meys"
date: "05/09/2019"
runtime: shiny
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Layout styles

## Recap on app structure

We've seen different kinds of apps :

 - **embedded in RMarkdown slides**
    * Only widgets needed
    * frontend/backend done for you
 - **single file apps in a file `app.R`**
    * Both UI and server in a single file
    * must return app object using `shinyApp()`
 - **two file apps: `ui.R` and `server.R`**
    * Must be in its own folder
    * `ui.R` returns a UI page object
    * `server.R` returns a function 

## Running an app.

Simplest way: clicking on a button in RStudio

More complex :

 - create `UI` and `server` objects
 - combine in a app object with `shinyApp()`
 - use `runApp()` to use the app.
 
```{r, eval = FALSE, echo = TRUE}
UIobj <- fluidPage( ... )
serverfun <- function(input, output){ ... }

theApp <- shinyApp(UIobj, serverfun)
runApp(theApp)
```

## Running an app.

Other ways to use `runApp()` :

 - using a list with a UI and server object:
```{r, eval = FALSE, echo = TRUE}
runApp(list(UI,serverfun))
```
 - Passing the directory where the files are located:
```{r, eval = FALSE, echo = TRUE}
runApp("path/to/appDir")
```

Allows us to create a function to call the app!

*See example `HistWithDens.R`*

## Layout principles of apps

Let's recap:

 - **Page**: the main container of an app, contains panels.
     * `fluidPage`
     * `shinydashboard`
 - **Panel** : a container containing other panels or widgets.
     * `sidebarPanel`
     * `mainPanel`
 - **Widget**: a functional unit (input, output)
     * `sliderInput`
     * `plotOutput`

This general principle goes for all different layout designs and possibilities.

## Fluid layout using `fluidPage()`

Fluid design is the standard in shiny. This means :

 - row based: a row is created using `fluidRow()`
     * They don't have a fixed height!
     * They arrange elements flowing left to right
 - Rows are divided in columns
     * there's 12 column units in a row
     * width of column in column units (1 - 12)

These are **panels** : can be nested in eachother

## Add HTML tags for formatting etc

`Shiny` contains a set of HTML functions :

 - `p()`: a paragraph with text
 - `br()`: a break
 - `a()`: a hyperlink
 - `h1()` to `h6()`: headers at different levels.
 
You can add attributes, eg:

```{r, echo=TRUE}
a("HTML tag reference!", href = "https://www.w3schools.com/TAGs/")
```

## Add navigation to the app

You can separate widgets in tabs using `tabPanel()`

Tabs have to be contained in :

 - `tabSetPanel()` : gives clickable tabs
 - `navlistPanel()` : gives navigation menu on the side
 - `navbarPage()` : a **page** with navigation bar for tabs
 
Keep in mind: `tabPanel()` is a container like `mainPanel()` etc.

## Control widget size :

Width and height can be set in CSS units for every widget.

 - is a character value with a number plus suffix.
 - suffices : `%`, `px`, `cm`, `mm`, `pt`, `in`, ...
 - relative width works relative to the column (eg `100%` = full column width)
 - relative height does NOT work! (height is determined by highest element.)
 
To use relative heights : 
 - use a `fillPage`
 - use a `fillRow` or `fillCol` with fixed height
 
