---
title: "Introduction to Shiny"
author: "Joris Meys"
date: "11/09/2020"
runtime: shiny
output: 
  ioslides_presentation:
    widescreen: true
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Profiling and Debugging

## Finding problems

Every code has problems, finding them is the trick.

### in normal R code:

 - use `print()` statements to see values
 - set breakpoints (either in the IDE or by using `browser()`)

### Problems in Shiny:

 - Using `print()` or `cat()` gives delayed results or might not work at all
 - breakpoints can only be set in RStudio IDE, other techniques won't work
 - breakpoints etc don't work on a server deployment (eg shinyapps.io)
 - No information about reactivity and dependency links.

## Classic debugging in Shiny :

### Printing

 - If you want to send messages, use
```{r, echo=TRUE, eval=FALSE}
cat(file = stderr(), "Input is", input$n,"\n Result is", res)
```
 - For more complex output, use
```{r, echo=TRUE, eval=FALSE}
capture.output({str(data)}, file = stderr())
```

By sending this to `stderr()`, you send the information through the channel
responsible for errors and warnings.

 - Doesn't interfere with the normal flow of information
 - is added to log files when app is run on a server

## Looking at reactions

[Reactlog](https://rstudio.github.io/reactlog/articles/reactlog.html) is a tool
that allows you to check how the reactivity works inside your app.

 - Install the package : `install.packages("reactlog")`
 - To initialize reactlog for shiny, use
```{r, echo = TRUE, eval = FALSE}
options(shiny.reactlog = TRUE)
```
 - run the app
 - after closing the app, use `reactlogShow()` to open the browser
 - to remove previous recordings, use `reactlogReset()`

## Why use reactlog?

 - Check the execution of the dependency chains.
 - Get an overview of the dependency links in your app
 - Identify areas/code that runs too often (or not at all)

## Looking at calculation time 

Reactlog gives you an idea, [ProfVis](https://rstudio.github.io/profvis/)
gives you a detailed look.

 - Can be used for all code, so not just in Shiny
 - to use with shiny app, simply run the app from within a call to `profvis`
 
```{r, echo = TRUE, eval = FALSE}
profvis(runApp("Apps/beaver1"))
```

# Modules

## Organizing your code

To re-use code in a shiny app: create a module.

 - reusable UI and server code to plug in a shiny app
 - ID tags are kept within the namespace of the module

### UI side

 - function taking `id` as first input (and possibly others)
 - it returns a `tagList` with UI elements (similar to `renderUI`)
 - it needs the namespace function (`ns <- NS(id)`)
 
### server side

 - function taking `input`, `output` and `session` as arguments
 - returns a reactive object 
 
## Calling the module

### UI side

 - simply use the ui function.
 - id you pass is the tag used to connect to the server side

```{r, eval=FALSE, echo=TRUE}
clickUI("tag", ...)
```

### server side

 - use `callModule` to call the function
 - return value is a reactive expression: catch it if needed
 
```{r, eval=FALSE, echo=TRUE}
dice1 <- callModule(clickServer, "tag")
```

**Remember to use `dice1` as a reactive, i.e. `dice1()` to access the value**

## Passing values to the server

The server side of a module can take extra arguments as well:

 - normal arguments just like an R function
 - reactive expressions: pass the object (no parentheses)
 
Rendering a UI at the server side requires the namespace function:

 - this is contained in the `session` object
 - extract using `ns <- session$ns`
 - Don't forget to use `ns(...)` in the rendered UI!

## The end

### Any questions?

Ask me anything.

### Exercise

Create a module for the beaver app that creates the ggplot and the slider.
Adapt the beaver app so you have a tab with the plot and range slider for
both beaver datasets. Use the module you created.

## Extra sources

 - Modules: http://shiny.rstudio.com/articles/modules.html
 - Modules2: http://shiny.rstudio.com/articles/communicate-bet-modules.html
 - reactlog: https://rstudio.github.io/reactlog/articles/reactlog.html
 - profvis: https://rstudio.github.io/profvis/
 - Debugging: http://shiny.rstudio.com/articles/debugging.html
