---
title: "Introduction to Shiny"
author: "Joris Meys"
date: "12/09/2019"
runtime: shiny
output: 
  ioslides_presentation:
    widescreen: true
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(miniUI)
```

# Sharing your applications

## Different goals require different solutions.

**Passing scripts**:

 - Helps R users
 - Users can change the code: flexible but unstable
 
**Integrating in RStudio with package**:

 - Helps R users
 - code is protected from unwanted changes
 - can be rolled out 
 
**Hosting on a web service**:

 - Also non-R users can use the app
 - Needs server deployment and admin
 - code is inaccessible/hidden

## Web services to host

Two main possibilities:

|        | Shiny Server (open) | ShinyProxy    |
| :----- | :----------------- | :----------   |
| From   | RStudio            | OpenAnalytics |
| App | Directory on server | Docker containers |
| Setup        | less complex   | more complex      |
| Users per app | All in same session | one session per user | 
| scalability  | Buy Pro        | in theory infinitely scaleable |

## Provide a GUI for complex functions

Until now: shiny apps as an independent unit: no return value

**Gadgets** :

 - Have a simple interface
 - Return a value when done
 - Are defined in a regular function
 - Can be used in a package or as RStudio addin
 
To build a gadget: need package `miniUI`

## Example of a gadget

```{r, echo = TRUE}
pickAcolor <- function(){
  
  ui <- miniPage(
    gadgetTitleBar("Select a color"),
    miniContentPanel(
      selectInput("col", "Pick a color",
                  choices = c("red","green","blue"))
    ))
  
  server <- function(input, output, session){
    observeEvent(input$done, {
      returnValue <- input$col
      stopApp(returnValue)
    })
  }
  runGadget(ui,server)
}
```

## Example of a gadget

 - Define `ui` and `server` using `miniUI`
 - Gadget always contains a `Cancel` and `Done` button.
 - use `observeEvent()` for detecting click on `Done`.
 - use `stopApp(val)` to return the value `val`
 - Last line in function : `runGadget`
 
To keep in mind :

 - `miniContentPanel()` is the default container to use.
 - Gadgets fill the window: use `fillRow` and `fillCol`, not `fluidRow` and `column`!
 - tabsets are handled by `miniUI` :
     * `miniTabstripPanel` for the container
     * `miniTabPanel` for the tabs itself
     
## Decorating Mini tabbed layout

Main use is the same as `tabPanel` and containers. 

 - `miniTabPanel` allows for icons using `icon` argument.
 - icons come from [FontAwesome](https://fontawesome.com/icons?d=gallery&m=free) and [Glyphicons](https://www.glyphicons.com/sets/basic/) iconset
 - Use function `icon()` to define them
 
```{r, echo = TRUE, eval = FALSE}
icon(name = "human", lib="glyphicon")
```
 
## Filled layouts

Gadgets use a flex box: tries to fill the entire page.

 - when using `plotOutput()`, set relative height

```{r, echo = TRUE, eval = FALSE}
miniContentPanel(
  plotOutput("theplot", height = "50%"),
  plotOutput("AnotherPlot", height = "50%")
)
```

 - To fit buttons/inputs on a row: `miniButtonBlock()`
 
```{r, echo = TRUE, eval = FALSE}
miniButtonBlock(
  actionButton("reset", "Reset"),
  actionButton("build", "build Model")
)
```

## Filled layouts

To use a button block in between the tabstrip and the tabs:

 - create a `miniButtonBlock()` object
 - pass to argument `between` of `miniTabstripPanel()`
 
```{r, echo = TRUE, eval = FALSE}
theButton <- miniButtonBlock(actionButton("reset","Reset"))
...
minTabstripPanel(
  between = theButton,
  miniTabPanel("Data", icon = icon("table"),
    ...
    ),
  miniTabPanel( ... )
  )
```

## Share online

 - Using Github gists:
     * Easy and "anonymous" sharing.
     * needs `ui.R` and `server.R` files in gist
     * run with `shiny::runGist('xxxxxxx')` where `xxxxxxx` is the gist ID number (you see that in the link)
 - Using Github repos:
     * Full github repo version control
     * Apps can be run from a subdirectory
     * run with `shiny::runGitHub()`

```{r, echo = TRUE, eval = FALSE}
runGitHub("Shiny_introduction",
          "JoFAM",
          subdir = "Class1_BasicApps/Apps/BasicPlot/")
```

## Share in a package

How to write a package : see eg R Packages by H.Wickham

For a shiny app:

 - place app in `inst/appdir`.
 - create (and export) a function that contains:
```{r, echo=TRUE,eval=FALSE}
shiny::runApp(system.file('appdir'),
              package = 'namePackage')
```
 - Make sure you declare dependency on Shiny!
 - App is started by calling the function.

## Add Gadgets/Shiny apps to RStudio

This has to be done using a package.

 - add Gadget function (or app function) to the package.
 - Register the app function in a `.dcf` file. (see example)
 - store `.dcf` file as `inst/rstudio/addins.dcf`
 
```
Name: The app
Description: This app does stuff
Binding: appFun
Interactive: true
```


## Extra sources

 - Shiny gadgets : http://shiny.rstudio.com/articles/gadgets.html
 - Gadget UI :http://shiny.rstudio.com/articles/gadget-ui.html
 - RStudio addins : http://rstudio.github.io/rstudioaddins/
 - Packages : http://r-pkgs.had.co.nz/
 - Local deployment : http://shiny.rstudio.com/articles/deployment-local.html

### Online deployment

 - ShinyProxy : https://www.shinyproxy.io/
 - ShinyServer setup reference : https://docs.rstudio.com/shiny-server/
